---
title: "Hands-on_Ex04"
author: "Vivian Fan Jing"
---

# Visual Statistical Analysis

# 10.2 Visual Statistical Analysis with ggstatsplot

[**ggstatsplot**](https://indrajeetpatil.github.io/ggstatsplot/index.html) is an extension of [**ggplot2**](https://ggplot2.tidyverse.org/) package for creating graphics with details from statistical tests included in the information-rich plots themselves.

-   To provide alternative statistical inference methods by default.
-   To follow best practices for statistical reporting. For all statistical tests reported in the plots, the default template abides by the [APA](https://my.ilstu.edu/~jhkahn/apastats.html) gold standard for statistical reporting.

# 10.3 Getting Started

## 10.3.1 Installing and launching R packages

In this exercise, ggstatsplot and tidyverse will be used.

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

## 10.3.2 Importing data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## 10.3.3 One-sample test: gghistostats() method

In the code chunk below, [*gghistostats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/gghistostats.html) is used to to build an visual of one-sample test on English scores.

```{r}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

## 10.3.6 Two-sample mean test: ggbetweenstats()

In the code chunk below, [*ggbetweenstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbetweenstats.html) is used to build a visual for two-sample mean test of Maths scores by gender

```{r}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

## 10.3.7 Oneway ANOVA Test: ggbetweenstats() method

In the code chunk below, [*ggbetweenstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbetweenstats.html) is used to build a visual for One-way ANOVA test on English score by race.

```{r}
ggbetweenstats(
  data = exam_data,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

-   “ns” → only non-significant

-   “s” → only significant

-   “all” → everything

## 10.3.8 Significant Test of Correlation: ggscatterstats()

In the code chunk below, [*ggscatterstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggscatterstats.html) is used to build a visual for Significant Test of Correlation between Maths scores and English scores.

```{r}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

## 10.3.9 Significant Test of Association (Depedence) : ggbarstats() methods

In the code chunk below, the Maths scores is binned into a 4-class variable by using [*cut()*](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cut).

In this code chunk below [*ggbarstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbarstats.html) is used to build a visual for Significant Test of Association

```{r}
exam1 <- exam_data %>%
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)

ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```
