---
title: "Take-home_Ex03_03"
author: "Vivian Fan Jing"
format: html
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
editor: visual
---

# 1. Installing and Launching R Packages

```{r}
pacman::p_load(SmartEDA, tidyverse, ggstatsplot,dplyr)
```

# 2 Data Wrangling

## 2.1 Importing the data set

In the code chunk below, read_csv() of readr is used to import WDI Country Data.csv into R and parsed it into tibble R data.frame format.

```{r}
water_indicators <- read.csv("WDI Country Data.csv")
```

## 2.2 Data summary

Function ExpData() comes from the SmartEDA package in R and The type = 2 argument specifies the type of analysis to perform.

```{r}
water_indicators %>%
  ExpData(type = 2)
```

## 2.3 Data Cleaning

```{r}
water_indicators[water_indicators == ".."] <- NA
water_indicators[water_indicators == "#N/A"] <- NA
water_indicators[water_indicators == ""] <- NA

```

```{r}
colnames(water_indicators) <- gsub("^X", "", colnames(water_indicators))
water_indicators_long <- water_indicators %>%
  pivot_longer(cols = starts_with("20"),  # Select columns starting with "20" (years)
               names_to = "year",         # New column for years
               values_to = "value")       # New column for values
```

Grouping  the years into categories and creating a new categorical column (year_group)

```{r}
water_indicators_long1 <- water_indicators_long %>%
  mutate(year = as.numeric(year)) %>%
  mutate(year_group = case_when(
    year >= 2000 & year <= 2005 ~ "2000-2005",
    year >= 2006 & year <= 2010 ~ "2006-2010",
    year >= 2011 & year <= 2015 ~ "2011-2015",
    year >= 2016 & year <= 2021 ~ "2016-2021",
    TRUE ~ NA_character_  # Handle unexpected years
  ))
```

# 3. Oneway ANOVA Test: ggbetweenstats() method over past Twenty years

ANOVA (Analysis of Variance) tests if the mean values of different groups are significantly different as we need statistical proof of differences, determining whether observed differences are statistically significant.

## 3.1 People using at least basic sanitation services (% of population) 

### 3.1.1 African Region

```{r}
water_long_selected1 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'African Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected1,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for African Region by Year Group"
)
```

### 3.1.2 Region of the Americas

```{r}
water_long_selected2 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'Region of the Americas')
```

```{r}
ggbetweenstats(
  data = water_long_selected2,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for Region of the Americas by Year Group"
)
```


### 3.1.3 European Region

```{r}
water_long_selected3 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'European Region')
```

```{r}
ggbetweenstats(
  data = water_long_selected3,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for European Region by Year Group"
)
```


### 3.1.4 Eastern Mediterranean Region

```{r}
water_long_selected4 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'Eastern Mediterranean Region')
```

```{r}
ggbetweenstats(
  data = water_long_selected4,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for Eastern Mediterranean Region by Year Group"
)
```


### 3.1.5 South-East Asia Region


```{r}
water_long_selected5 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'South-East Asia Region')
```

```{r}
ggbetweenstats(
  data = water_long_selected5,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for South-East Asia Region by Year Group"
)
```


### 3.1.6 Western Pacific Region


```{r}
water_long_selected6 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'Western Pacific Region')
```

```{r}
ggbetweenstats(
  data = water_long_selected6,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for Western Pacific Region by Year Group"
)
```

## 3.2 People using safely managed sanitation services (% of population)

### 3.2.1 African Region

```{r}
water_long_selected7 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'African Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected7,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for African Region by Year Group")
```

### 3.2.2 Region of the Americas


```{r}
water_long_selected8 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'Region of the Americas')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected8,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for Region of the Americas"
)
```



### 3.2.3 European Region

```{r}
water_long_selected9 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'European Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected9,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for European Region"
)
```



### 3.2.4 Eastern Mediterranean Region

```{r}
water_long_selected10 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'Eastern Mediterranean Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected10,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for Eastern Mediterranean Region"
)
```


### 3.2.5 South-East Asia Region


```{r}
water_long_selected11 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'South-East Asia Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected11,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for South-East Asia Region"
)
```

### 3.2.6 Western Pacific Region


```{r}
water_long_selected12 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'Western Pacific Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected12,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for Western Pacific Region"
)
```

## 3.3 People using at least basic drinking water services (% of population)

### 3.3.1 African Region

```{r}
water_long_selected13 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'African Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected13,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for African Region by Year Group"
)
```


### 3.3.2 Region of the Americas


```{r}
water_long_selected14 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'Region of the Americas')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected14,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Region of the Americas"
)
```



### 3.3.3 European Region

```{r}
water_long_selected15 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'European Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected15,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for European Region"
)
```


### 3.3.4 Eastern Mediterranean Region


```{r}
water_long_selected16 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'Eastern Mediterranean Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected16,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Eastern Mediterranean Region"
)
```



### 3.3.5 South-East Asia Region


```{r}
water_long_selected17 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'South-East Asia Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected17,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for South-East Asia Region"
)
```

### 3.3.6 Western Pacific Region


```{r}
water_long_selected18 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'Western Pacific Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected18,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Western Pacific Region"
)
```

## 3.4 People using safely managed drinking water services (% of population)

### 3.4.1 African Region


```{r}
water_long_selected19 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'African Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected19,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for African Region by Year Group"
)
```


### 3.4.2 Region of the Americas


```{r}
water_long_selected20 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'Region of the Americas')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected20,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Region of the Americas by Year Group"
)
```


### 3.4.3 European Region


```{r}
water_long_selected21 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'European Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected21,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for European Region by Year Group"
)
```

### 3.4.4 Eastern Mediterranean Region


```{r}
water_long_selected22 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'Eastern Mediterranean Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected22,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Eastern Mediterranean Region"
)
```

### 3.4.5 South-East Asia Region


```{r}
water_long_selected23 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'South-East Asia Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected23,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for South-East Asia Region"
)
```

### 3.4.6 Western Pacific Region


```{r}
water_long_selected24 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'Western Pacific Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected24,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Western Pacific Region by Year Group"
)
```



## 3.5 People practicing open defecation (% of population)

### 3.5.1 African Region

```{r}
water_long_selected25 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'African Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected25,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for African Region by Year Group"
)
```

### 3.5.2 Region of the Americas


```{r}
water_long_selected26 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'Region of the Americas')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected26,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for Region of the Americas"
)
```


### 3.5.3 European Region

```{r}
water_long_selected27 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'European Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected27,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for European Region"
)
```

### 3.5.4 Eastern Mediterranean Region


```{r}
water_long_selected28 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'Eastern Mediterranean Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected28,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for Eastern Mediterranean Region"
)
```

### 3.5.5 South-East Asia Region

```{r}
water_long_selected29 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'South-East Asia Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected29,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for South-East Asia Region"
)
```

### 3.5.6 Western Pacific Region


```{r}
water_long_selected30 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'Western Pacific Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected30,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for Western Pacific Region"
)
```

## 3.6 Level of water stress: freshwater withdrawal as a proportion of available freshwater resources

### 3.6.1 African Region

```{r}
water_long_selected31 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'African Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected31,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for African Region by Year Group"
)
```



### 3.6.2 Region of the Americas


```{r}
water_long_selected32 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'Region of the Americas')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected32,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for Region of the Americas"
)
```
### 3.6.3 European Region

```{r}
water_long_selected33 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'European Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected33,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for European Region"
)
```

### 3.6.4 Eastern Mediterranean Region

```{r}
water_long_selected34 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'Eastern Mediterranean Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected34,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for Eastern Mediterranean Region"
)
```

### 3.6.5 South-East Asia Region


```{r}
water_long_selected35 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'South-East Asia Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected35,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for South-East Asia Region"
)
```


### 3.6.6 Western Pacific Region

```{r}
water_long_selected36 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'Western Pacific Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected36,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for Western Pacific Region"
)
```

## 3.7 People with basic handwashing facilities including soap and water (% of population)

### 3.7.1 African Region


```{r}
water_long_selected37 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'African Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected37,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for African Region by Year Group"
)
```


### 3.7.2 Region of the Americas

```{r}
water_long_selected38 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'Region of the Americas')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected38,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for Region of the Americas"
)
```



### 3.7.3 European Region


```{r}
water_long_selected39 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'European Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected39,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for European Region"
)
```


### 3.7.4 Eastern Mediterranean Region


```{r}
water_long_selected40 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'Eastern Mediterranean Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected40,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for Eastern Mediterranean Region"
)
```

### 3.7.5 South-East Asia Region


```{r}
water_long_selected41 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'South-East Asia Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected41,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for South-East Asia Region"
)
```

### 3.7.6 Western Pacific Region


```{r}
water_long_selected42 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'Western Pacific Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected42,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for Western Pacific Region"
)
```


## 3.8 Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)

### 3.8.1 African Region

```{r}
water_long_selected43 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'African Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected43,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for African Region by Year Group"
)
```


### 3.8.2 Region of the Americas


```{r}
water_long_selected44 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'Region of the Americas')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected44,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for Region of the Americas"
)
```


### 3.8.3 European Region

```{r}
water_long_selected45 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'European Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected45,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for European Region"
)
```
### 3.8.4 Eastern Mediterranean Region


```{r}
water_long_selected46 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'Eastern Mediterranean Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected46,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for Eastern Mediterranean Region"
)
```

### 3.8.5 South-East Asia Region

```{r}
water_long_selected47 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'South-East Asia Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected47,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for South-East Asia Region"
)
```

### 3.8.6 Western Pacific Region

```{r}
water_long_selected48 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'Western Pacific Region')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected48,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for Western Pacific Region"
)
```
