---
title: "Take-home_Ex03_03"
author: "Vivian Fan Jing"
format: html
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
editor: visual
---

# 1. Installing and Launching R Packages

```{r}
pacman::p_load(SmartEDA, tidyverse, ggstatsplot,dplyr)
```

# 2 Data Wrangling

## 2.1 Importing the data set

In the code chunk below, read_csv() of readr is used to import WDI Country Data.csv into R and parsed it into tibble R data.frame format.

```{r}
water_indicators <- read.csv("WDI Country Data.csv")
```

## 2.2 Data summary

Function ExpData() comes from the SmartEDA package in R and The type = 2 argument specifies the type of analysis to perform.

```{r}
water_indicators %>%
  ExpData(type = 2)
```

## 2.3 Data Cleaning

```{r}
water_indicators[water_indicators == ".."] <- NA
water_indicators[water_indicators == "#N/A"] <- NA
water_indicators[water_indicators == ""] <- NA

```

```{r}
colnames(water_indicators) <- gsub("^X", "", colnames(water_indicators))
water_indicators_long <- water_indicators %>%
  pivot_longer(cols = starts_with("20"),  # Select columns starting with "20" (years)
               names_to = "year",         # New column for years
               values_to = "value")       # New column for values
```

Grouping  the years into categories and creating a new categorical column (year_group)

```{r}
water_indicators_long1 <- water_indicators_long %>%
  mutate(year = as.numeric(year)) %>%
  mutate(year_group = case_when(
    year >= 2000 & year <= 2005 ~ "2000-2005",
    year >= 2006 & year <= 2010 ~ "2006-2010",
    year >= 2011 & year <= 2015 ~ "2011-2015",
    year >= 2016 & year <= 2021 ~ "2016-2021",
    TRUE ~ NA_character_  # Handle unexpected years
  ))
```

# 3. Oneway ANOVA Test: ggbetweenstats() method over past Twenty years

ANOVA (Analysis of Variance) tests if the mean values of different groups are significantly different as we need statistical proof of differences, determining whether observed differences are statistically significant.

## 3.1 People using at least basic sanitation services (% of population) 

### 3.1.1 Sub-Saharan Africa

```{r}
water_long_selected1 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'Sub-Saharan Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected1,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for Sub-Saharan Africa by Year Group"
)
```

### 3.1.2 Latin America & Caribbean

```{r}
water_long_selected2 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'Latin America & Caribbean')
```

```{r}
ggbetweenstats(
  data = water_long_selected2,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for Latin America & Caribbean by Year Group"
)
```


### 3.1.3 Europe & Central Asia

```{r}
water_long_selected3 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'Europe & Central Asia')
```

```{r}
ggbetweenstats(
  data = water_long_selected3,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for Europe & Central Asia by Year Group"
)
```


### 3.1.4 Middle East & North Africa

```{r}
water_long_selected4 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'Middle East & North Africa')
```

```{r}
ggbetweenstats(
  data = water_long_selected4,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for Middle East & North Africa by Year Group"
)
```


### 3.1.5 East Asia & Pacific


```{r}
water_long_selected5 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'East Asia & Pacific')
```

```{r}
ggbetweenstats(
  data = water_long_selected5,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for East Asia & Pacific by Year Group"
)
```


### 3.1.6 South Asia


```{r}
water_long_selected6 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'South Asia')
```

```{r}
ggbetweenstats(
  data = water_long_selected6,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for South Asia by Year Group"
)
```


### 3.1.7 North America


```{r}
water_long_selected49 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic sanitation services (% of population)',
         Region == 'North America')
```

```{r}
ggbetweenstats(
  data = water_long_selected49,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Sanitation Services for North America by Year Group"
)
```

## 3.2 People using safely managed sanitation services (% of population)

### 3.2.1 Sub-Saharan Africa

```{r}
water_long_selected7 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'Sub-Saharan Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected7,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for Sub-Saharan Africa by Year Group")
```

### 3.2.2 Latin America & Caribbean


```{r}
water_long_selected8 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'Latin America & Caribbean')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected8,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for Latin America & Caribbean"
)
```



### 3.2.3 Europe & Central Asia

```{r}
water_long_selected9 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'Europe & Central Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected9,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for Europe & Central Asia"
)
```



### 3.2.4 Middle East & North Africa

```{r}
water_long_selected10 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'Middle East & North Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected10,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for Middle East & North Africa"
)
```


### 3.2.5 East Asia & Pacific


```{r}
water_long_selected11 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'East Asia & Pacific')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected11,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for East Asia & Pacific"
)
```

### 3.2.6 South Asia


```{r}
water_long_selected12 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'South Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected12,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for South Asia by Year Group"
)
```

### 3.2.7 North America


```{r}
water_long_selected50 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed sanitation services (% of population)',
         Region == 'North America')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected50,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Sanitation Services for North America by Year Group"
)
```


## 3.3 People using at least basic drinking water services (% of population)

### 3.3.1 Sub-Saharan Africa

```{r}
water_long_selected13 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'Sub-Saharan Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected13,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Sub-Saharan Africa by Year Group"
)
```


### 3.3.2 Latin America & Caribbean


```{r}
water_long_selected14 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'Latin America & Caribbean')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected14,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Latin America & Caribbean"
)
```



### 3.3.3 Europe & Central Asia

```{r}
water_long_selected15 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'Europe & Central Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected15,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Europe & Central Asia"
)
```


### 3.3.4 Middle East & North Africa


```{r}
water_long_selected16 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'Middle East & North Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected16,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for Middle East & North Africa"
)
```



### 3.3.5 East Asia & Pacific


```{r}
water_long_selected17 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'East Asia & Pacific')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected17,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for East Asia & Pacific by Year Group"
)
```

### 3.3.6 South Asia


```{r}
water_long_selected18 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'South Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected18,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for South Asia by Year Group"
)
```

### 3.3.7 North America


```{r}
water_long_selected51 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using at least basic drinking water services (% of population)',
         Region == 'North America')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected51,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Drinking Water Services for North America by Year Group"
)
```



## 3.4 People using safely managed drinking water services (% of population)

### 3.4.1 Sub-Saharan Africa


```{r}
water_long_selected19 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'Sub-Saharan Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected19,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Drinking Water Services for Sub-Saharan Africa "
)
```


### 3.4.2 Latin America & Caribbean


```{r}
water_long_selected20 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'Latin America & Caribbean')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected20,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Drinking Water Services for Latin America & Caribbean by Year Group"
)
```


### 3.4.3 Europe & Central Asia


```{r}
water_long_selected21 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'Europe & Central Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected21,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Drinking Water Services for Europe & Central Asia"
)
```

### 3.4.4 Middle East & North Africa


```{r}
water_long_selected22 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'Middle East & North Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected22,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Drinking Water Services for Middle East & North Africa"
)
```

### 3.4.5 East Asia & Pacific


```{r}
water_long_selected23 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'East Asia & Pacific')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected23,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Drinking Water Services for East Asia & Pacific"
)
```

### 3.4.6 South Asia


```{r}
water_long_selected24 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'South Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected24,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Drinking Water Services for South Asia by Year Group"
)
```


### 3.4.7 North America


```{r}
water_long_selected52 <- water_indicators_long1 %>%
  filter(Series.Name == 'People using safely managed drinking water services (% of population)',
         Region == 'North America')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected52,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Safely Managed Drinking Water Services for North America by Year Group"
)
```

## 3.5 People practicing open defecation (% of population)

### 3.5.1 Sub-Saharan Africa

```{r}
water_long_selected25 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'Sub-Saharan Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected25,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for Sub-Saharan Africa by Year Group"
)
```

### 3.5.2 Latin America & Caribbean


```{r}
water_long_selected26 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'Latin America & Caribbean')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected26,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for Latin America & Caribbean"
)
```


### 3.5.3 Europe & Central Asia

```{r}
water_long_selected27 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'Europe & Central Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected27,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for Europe & Central Asia"
)
```

### 3.5.4 Middle East & North Africa


```{r}
water_long_selected28 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'Middle East & North Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected28,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for Middle East & North Africa"
)
```

### 3.5.5 East Asia & Pacific

```{r}
water_long_selected29 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'East Asia & Pacific')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected29,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for East Asia & Pacific by Year Group"
)
```

### 3.5.6 South Asia


```{r}
water_long_selected30 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'South Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected30,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for South Asia"
)
```

### 3.5.7 North America


```{r}
water_long_selected53 <- water_indicators_long1 %>%
  filter(Series.Name == 'People practicing open defecation (% of population)',
         Region == 'South Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected53,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Practicing Open Defecation for South Asia"
)
```



## 3.6 Level of water stress: freshwater withdrawal as a proportion of available freshwater resources

### 3.6.1 Sub-Saharan Africa

```{r}
water_long_selected31 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'Sub-Saharan Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected31,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for Sub-Saharan Africa by Year Group"
)
```



### 3.6.2 Latin America & Caribbean


```{r}
water_long_selected32 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'Latin America & Caribbean')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected32,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for Latin America & Caribbean"
)
```
### 3.6.3 Europe & Central Asia

```{r}
water_long_selected33 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'Europe & Central Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected33,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for Europe & Central Asia"
)
```

### 3.6.4 Middle East & North Africa

```{r}
water_long_selected34 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'Middle East & North Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected34,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for Middle East & North Africa"
)
```

### 3.6.5 East Asia & Pacific

```{r}
water_long_selected35 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'East Asia & Pacific')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected35,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for East Asia & Pacific by Year Group"
)
```


### 3.6.6 South Asia

```{r}
water_long_selected36 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'South Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected36,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for South Asia by Year Group"
)
```



### 3.6.7 North America


```{r}
water_long_selected54 <- water_indicators_long1 %>%
  filter(Series.Name == 'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
         Region == 'North America')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected54,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Level of water stress for North America by Year Group"
)
```



## 3.7 People with basic handwashing facilities including soap and water (% of population)

### 3.7.1 Sub-Saharan Africa


```{r}
water_long_selected37 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'Sub-Saharan Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected37,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for Sub-Saharan Africa by Year Group"
)
```


### 3.7.2 Latin America & Caribbean

```{r}
water_long_selected38 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'Latin America & Caribbean')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected38,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for Latin America & Caribbean"
)
```



### 3.7.3 Europe & Central Asia


```{r}
water_long_selected39 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'Europe & Central Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected39,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for Europe & Central Asia"
)
```


### 3.7.4 Middle East & North Africa


```{r}
water_long_selected40 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'Middle East & North Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected40,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for Middle East & North Africa"
)
```

### 3.7.5 East Asia & Pacific


```{r}
water_long_selected41 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'East Asia & Pacific')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected41,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for East Asia & Pacific"
)
```

### 3.7.6 South Asia


```{r}
water_long_selected42 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'South Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected42,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for South Asia by Year Group"
)
```

### 3.7.7 North America


```{r}
water_long_selected55 <- water_indicators_long1 %>%
  filter(Series.Name == 'People with basic handwashing facilities including soap and water (% of population)',
         Region == 'North America')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected55,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Basic Handwashing Facilities for North America"
)
```


## 3.8 Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)

### 3.8.1 Sub-Saharan Africa

```{r}
water_long_selected43 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'Sub-Saharan Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected43,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for Sub-Saharan Africa by Year Group"
)
```


### 3.8.2 Latin America & Caribbean


```{r}
water_long_selected44 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'Latin America & Caribbean')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected44,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for Latin America & Caribbean by Year Group"
)
```


### 3.8.3 Europe & Central Asia

```{r}
water_long_selected45 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'Europe & Central Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected45,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for Europe & Central Asia by Year Group"
)
```
### 3.8.4 Middle East & North Africa


```{r}
water_long_selected46 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'Middle East & North Africa')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected46,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for Middle East & North Africa by Year Group"
)
```

### 3.8.5 East Asia & Pacific

```{r}
water_long_selected47 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'East Asia & Pacific')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected47,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for East Asia & Pacific by Year Group"
)
```

### 3.8.6 South Asia

```{r}
water_long_selected48 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'South Asia')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected48,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for South Asia by Year Group"
)
```
### 3.8.7 North America



```{r}
water_long_selected56 <- water_indicators_long1 %>%
  filter(Series.Name == 'Water productivity, total (constant 2015 US$ GDP per cubic meter of total freshwater withdrawal)',
         Region == 'North America')
           
```

```{r}
ggbetweenstats(
  data = water_long_selected56,
  x = year_group, 
  y = value, 
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "significant", 
  p.adjust.method = "fdr",
  messages = FALSE,
  title = "Comparison of Water Productivity for North America by Year Group"
)
```

