---
title: "Hands-on_Ex05_03"
author: "Vivian Fan Jing"
date: "February 12, 2025"
execute:
  warning: false
---

# 15 Visual Multivariate Analysis with Parallel Coordinates Plot

The strength of parallel coordinates isn’t in their ability to communicate some truth in the data to others, but rather in their ability to bring meaningful multivariate patterns and comparisons to light when used interactively for analysis.

-   plotting statistic parallel coordinates plots by using *ggparcoord()* of **GGally** package,

-   plotting interactive parallel coordinates plots by using **parcoords** package, and

-   plotting interactive parallel coordinates plots by using **parallelPlot** package.

## 15.2 Installing and Launching R Packages

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

## 15.4 Plotting Static Parallel Coordinates Plot

plot static parallel coordinates plot by using [`ggparcoord()`](https://ggobi.github.io/ggally/reference/ggparcoord.html) of **GGally** package

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

### 15.4.2 Plotting a parallel coordinates with boxplot

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

Things to learn from the code chunk above.

-   `groupColumn` argument is used to group the observations (i.e. parallel lines) by using a single variable (i.e. Region) and colour the parallel coordinates lines by region name.

-   `scale` argument is used to scale the variables in the parallel coordinate plot by using `uniminmax` method. The method univariately scale each variable so the minimum of the variable is zero and the maximum is one.

-   `alphaLines` argument is used to reduce the intensity of the line colour to 0.2. The permissible value range is between 0 to 1.

-   `boxplot` argument is used to turn on the boxplot by using logical `TRUE`. The default is `FALSE`.

-   `title` argument is used to provide the parallel coordinates plot a title.

### 15.4.3 Parallel coordinates with facet

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)
```

One of the aesthetic defect of the current design is that some of the variable names overlap on x-axis.

### 15.4.4 Rotating x-axis text label

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))
```

Thing to learn from the code chunk above:

-   To rotate x-axis text labels, we use `axis.text.x` as argument to `theme()` function. And we specify `element_text(angle = 30)` to rotate the x-axis text by an angle 30 degree.

### 15.4.5 Adjusting the rotated x-axis text label

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

## 15.5 Plotting Interactive Parallel Coordinates Plot: parallelPlot methods

[**parallelPlot**](https://cran.r-project.org/web/packages/parallelPlot/) is an R package specially designed to plot a parallel coordinates plot by using ‘htmlwidgets’ package and [d3.js](https://d3js.org/). In this section, you will learn how to use functions provided in parallelPlot package to build interactive parallel coordinates plot.

### 15.5.1 The basic plot

```{r}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)
```

### 15.5.2 Rotate axis label

```{r}
parallelPlot(wh,
             rotateTitle = TRUE)
```

One of the useful interactive feature of parallelPlot is we can click on a variable of interest, for example Happiness score, the monotonous blue colour (default) will change a blues with different intensity colour scheme will be used.

### 15.5.3 Changing the colour scheme

We can change the default blue colour scheme by using `continousCS` argument as shown in the code chunl below.

```{r}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

### 15.5.4 Parallel coordinates plot with histogram

In the code chunk below, `histoVisibility` argument is used to plot histogram along the axis of each variables.

```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```

## 15.6 References

-   [*ggparcoord()*](http://ggobi.github.io/ggally/reference/ggparcoord.html) of [**GGally**](http://ggobi.github.io/ggally/index.html) package

-   [**parcoords** user guide](https://cran.r-project.org/web/packages/parcoords/parcoords.pdf)

-   [**parallelPlot**](https://cran.r-project.org/web/packages/parallelPlot/parallelPlot.pdf)
