---
title: "Hands-on_Ex08"
author: "Vivian Fan Jing"
date: "March 13, 2025"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Modelling, Visualising and Analysing Network Data with R

## Installing and launching R packages

the key R package use is [**tmap**](https://cran.r-project.org/web/packages/tmap/) package in R. Beside **tmap** package, four other R packages will be used. They are:

-   [**readr**](https://readr.tidyverse.org/) for importing delimited text file,

-   [**tidyr**](https://tidyr.tidyverse.org/) for tidying data,

-   [**dplyr**](https://dplyr.tidyverse.org/) for wrangling data and

-   [**sf**](https://cran.r-project.org/web/packages/sf/index.html) for handling geospatial data.

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## The Data

Two data set will be used to create the choropleth map. They are:

-   Master Plan 2014 Subzone Boundary (Web)

-   Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format

### Importing Geospatial Data into R

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

Examining the content of mpsz by using the code chunk below

```{r}
mpsz
```

### Importing Attribute Data into R

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

### Data Preparation

Before a thematic map can be prepared, you are required to prepare a data table with year 2020 values. The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: age group 0 to 4 until age groyup 20 to 24,

-   ECONOMY ACTIVE: age group 25-29 until age group 60-64,

-   AGED: age group 65 and above,

-   TOTAL: all age group, and

-   DEPENDENCY: the ratio between young and aged against economy active group

**Data wrangling**

The following data wrangling and transformation functions will be used:

-   *pivot_wider()* of **tidyr** package, and

-   *mutate()*, *filter()*, *group_by()* and *select()* of **dplyr** package

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

**Joining the attribute data and geospatial data**

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

*left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## Choropleth Mapping Geospatial Data Using tmap

Two approaches can be used to prepare thematic map using *tmap*, they are:

-   Plotting a thematic map quickly by using *qtm()*.

-   Plotting highly customisable thematic map by using tmap elements.

### Plotting a choropleth map quickly by using qtm()

The easiest and quickest to draw a choropleth map using **tmap** is using *qtm()*. It is concise and provides a good default visualisation in many cases.

The code chunk below will draw a cartographic standard choropleth map as shown below.

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

Things to learn from the code chunk above:

-   *tmap_mode()* with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.

-   *fill* argument is used to map the attribute (i.e. DEPENDENCY)

### Creating a choropleth map by using tmap’s elements

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

#### 21.4.2.1 Drawing a base map

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### Drawing a choropleth map using tm_polygons()

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

#### 21.4.2.3 Drawing a choropleth map using tm_fill() and \*tm_border()\*\*

Actually, tm_polygons() is a wraper of tm_fill() and tm_border(). tm_fill() shades the polygons by using the default colour scheme and tm_borders() adds the borders of the shapefile onto the choropleth map.

The code chunk below draws a choropleth map by using tm_fill() alone.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

Notice that the planning subzones are shared according to the respective dependecy values

To add the boundary of the planning subzones, tm_borders will be used as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```

Notice that light-gray border lines have been added on the choropleth map.

The *alpha* argument is used to define transparency number between 0 (totally transparent) and 1 (not transparent). By default, the alpha value of the col is used (normally 1).

Beside *alpha* argument, there are three other arguments for *tm_borders()*, they are:

-   *col* = border colour,

-   *lwd* = border line width. The default is 1, and

-   *lty* = border line type. The default is “solid”.

### Data classification methods of tmap

#### Plotting choropleth maps with built-in classification methods

The code chunk below shows a quantile data classification that used 5 classes.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

#### Plotting choropleth map with custome break

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

### Colour Scheme

#### 21.4.4.1 Using ColourBrewer palette

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

Notice that the choropleth map is shaded in green.

To reverse the colour shading, add a “-” prefix.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

### Map Layouts

#### Map Legend

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

#### Map style

tmap allows a wide variety of layout settings to be changed.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

##### Cartographic Furniture

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

To reset the default style, refer to the code chunk below.

```{r}
tmap_style("white")
```

### Drawing Small Multiple Choropleth Maps

#### By assigning multiple values to at least one of the aesthetic arguments

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

In this example, small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments

```{r}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

#### By defining a group-by variable in tm_facets()

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

#### By creating multiple stand-alone maps with tmap_arrange()

In this example, multiple small choropleth maps are created by creating multiple stand-alone maps with **tmap_arrange()**.

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### Mappping Spatial Object Meeting a Selection Criterion

Instead of creating small multiple choropleth map, you can also use selection funtion to map spatial objects meeting the selection criterion.

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

