---
title: "Hands-on_Ex07"
author: "Vivian Fan Jing"
date: "February 26, 2025"
execute:
  warning: false
---

# 17 Visualising and Analysing Time-oriented Data

Creating data visualization by using R packages:

-   plotting a calender heatmap by using ggplot2 functions,

-   plotting a cycle plot by using ggplot2 function,

-   plotting a slopegraph

-   plotting a horizon chart

Write a code chunk to check, install and launch the following R packages: scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table and tidyverse.

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table, CGPfunctions, ggHoriPlot, tidyverse)
```

## 17.4 Plotting Calendar Heatmap

plot a calender heatmap programmatically by using ggplot2 package

### 17.4.1 Importing the data

In this hands-on exercise, eventlog.csv file will be used. This data file consists of 199,999 rows of time-series cyber attack records by country.

```{r}
attacks <- read_csv("data/eventlog.csv")
```

### 17.4.2 Examining the data structure

kable() is used to review the structure of the imported data frame.

```{r}
kable(head(attacks))
```

There are three columns, namely timestamp, source_country and tz.

There are three columns, namely *timestamp*, *source_country* and *tz*.

-   *timestamp* field stores date-time values in POSIXct format.

-   *source_country* field stores the source of the attack. It is in *ISO 3166-1 alpha-2* country code.

-   *tz* field stores time zone of the source IP address.

### 17.4.3 Data Preparation

Step 1: Deriving weekday and hour of day fields

Two new fields namely wkday and hour need to be derived. In this step, we will write a function to perform the task.

```{r}
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
  }
```

-   [`ymd_hms()`](https://lubridate.tidyverse.org/reference/ymd_hms.html) and [`hour()`](https://lubridate.tidyverse.org/reference/hour.html) are from [**lubridate**](https://lubridate.tidyverse.org/) package, and

-   [`weekdays()`](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/weekdays) is a **base** R function.

Step 2: Deriving the attacks tibble data frame

wkday_levels \<- c('Saturday', 'Friday', 'Thursday', 'Wednesday', 'Tuesday', 'Monday', 'Sunday')

```{r}

wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))
```

Table below shows the tidy tibble table after processing.

```{r}
kable(head(attacks))
```

### 17.4.4 Building the Calendar Heatmaps

```{r}
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

-   a tibble data table called *grouped* is derived by aggregating the attack by *wkday* and *hour* fields.

-   a new field called *n* is derived by using `group_by()` and `count()` functions.

-   `na.omit()` is used to exclude missing value.

-   `geom_tile()` is used to plot tiles (grids) at each x and y position. `color` and `size` arguments are used to specify the border color and line size of the tiles.

-   [`theme_tufte()`](https://jrnold.github.io/ggthemes/reference/theme_tufte.html) of [**ggthemes**](https://jrnold.github.io/ggthemes/reference/index.html) package is used to remove unnecessary chart junk. To learn which visual components of default ggplot2 have been excluded, you are encouraged to comment out this line to examine the default plot.

-   `coord_equal()` is used to ensure the plot will have an aspect ratio of 1:1.

-   `scale_fill_gradient()` function is used to creates a two colour gradient (low-high).

### 17.4.5 Building Multiple Calendar Heatmaps

Step 1: Deriving attack by country object

-   count the number of attacks by country,

-   calculate the percent of attackes by country, and

-   save the results in a tibble data frame.

```{r}
attacks_by_country <- count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))
```


Step 2: Preparing the tidy data frame

Extracting the attack records of the top 4 countries from attacks data frame and saving the data in a new tibble data frame (i.e. top4_attacks).

```{r}
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
```


Step 3: Plotting the Multiple Calender Heatmap by using ggplot2 package

```{r}
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## 17.4 Plotting Cycle Plot

plot a cycle plot showing the time-series patterns and trend of visitor arrivals from Vietnam programmatically by using ggplot2 functions.

Step 1: Data Import

```{r}
air <- read_excel("data/arrivals_by_air.xlsx")
```

Step 2: Deriving month and year fields

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

Step 3: Extracting the target country

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

Step 4: Computing year average arrivals by month

```{r}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

Step 5: Plotting the cycle plot

```{r}
ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")
```


## 17.5 Plotting Slopegraph

Step 1: Data Import

```{r}
rice <- read_csv("data/rice.csv")
```

Step 2: Plotting the slopegraph

```{r}
rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  newggslopegraph(Year, Yield, Country,
                Title = "Rice Yield of Top 11 Asian Counties",
                SubTitle = "1961-1980",
                Caption = "Prepared by: Vivian Fan Jing")
```


