---
title: "Hands-on_Ex08_01"
author: "Vivian Fan Jing"
date: "March 13, 2025"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Visualising Geospatial Point Data

In this hands-on exercise, I acquire the following skills by using appropriate R packages:

-   To import an aspatial data file into R.

-   To convert it into simple point feature data frame and at the same time, to assign an appropriate projection reference to the newly create simple point feature data frame.

-   To plot interactive proportional symbol maps.

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## The data

The data set use for this hands-on exercise is called SGPools_svy21. The data is in csv file format.

## Data Import and Preparation

The code chunk below uses *read_csv()* function of **readr** package to import *SGPools_svy21.csv* into R as a tibble data frame called *sgpools*.

```{r}
sgpools <- read_csv("data/SGPools_svy21.csv")
```

The code chunk below shows list() is used to do the job.

```{r}
list(sgpools)
```

Notice that the sgpools data in tibble data frame and not the common R data frame.

## Creating a sf data frame from an aspatial data frame

The code chunk below converts sgpools data frame into a simple feature data frame by using st_as_sf() of sf packages

```{r}
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c("XCOORD", "YCOORD"),
                       crs= 3414)
```

Things to learn from the arguments above:

-   The *coords* argument requires you to provide the column name of the x-coordinates first then followed by the column name of the y-coordinates.

-   The *crs* argument required you to provide the coordinates system in epsg format. [EPSG: 3414](https://epsg.io/3414) is Singapore SVY21 Projected Coordinate System. You can search for other country’s epsg code by refering to [epsg.io](https://epsg.io/).

display the basic information of the newly created sgpools_sf by using the code chunk below

```{r}
list(sgpools_sf)
```

## It all started with an interactive point symbol map

The code chunks below are used to create an interactive point symbol map.

```{r}
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = 1,
           border.col = "black",
           border.lwd = 1)
```

## Lets make it proportional

To draw a proportional symbol map, we need to assign a numerical variable to the size visual attribute. The code chunks below show that the variable Gp1Gp2Winnings is assigned to size visual attribute.

```{r}
tm_shape(sgpools_sf)+
tm_bubbles(col = "red",
           size = "Gp1Gp2 Winnings",
           border.col = "black",
           border.lwd = 1)
```

## Lets give it a different colour

The proportional symbol map can be further improved by using the colour visual attribute. In the code chunks below, OUTLET_TYPE variable is used as the colour attribute variable.

```{r}
tm_shape(sgpools_sf)+
tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1)
```
## I have a twin brothers :)

An impressive and little-know feature of tmap’s view mode is that it also works with faceted plots. The argument sync in tm_facets() can be used in this case to produce multiple maps with synchronised zoom and pan settings.

```{r}
tm_shape(sgpools_sf) +
  tm_bubbles(col = "OUTLET TYPE", 
          size = "Gp1Gp2 Winnings",
          border.col = "black",
          border.lwd = 1) +
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```

switch tmap’s Viewer back to plot mode by using the code chunk below.

 
 