---
title: "Hands-on_Ex09"
author: "Vivian Fan Jing"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Modelling, Visualising and Analysing Network Data with R

In this hands-on exercise:

-   create graph object data frames, manipulate them using appropriate functions of *dplyr*, *lubridate*, and *tidygraph*,

-   build network graph visualisation using appropriate functions of *ggraph*,

-   compute network geometrics using *tidygraph*,

-   build advanced graph visualisation by incorporating the network geometrics, and

-   build interactive network visualisation using *visNetwork* package.

## Getting Started

### Installing and launching R packages

```{r}
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts, 
               concaveman, ggforce)
```

## The Data

The data sets used in this hands-on exercise is from an oil exploration and extraction company. There are two data sets. One contains the nodes data and the other contains the edges (also know as link) data.

-   GAStech-email_edges.csv which consists of two weeks of 9063 emails correspondances between 55 employees.

-   GAStech_email_nodes.csv which consist of the names, department and title of the 55 employees.

### Importing network data from files

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")
```

### Reviewing the imported data

Examining the structure of the data frame using *glimpse()* of **dplyr**.

```{r}
glimpse(GAStech_edges)
```

### Wrangling time

The code chunk below will be used to perform the changes.

```{r}
GAStech_edges <- GAStech_edges %>%
  mutate(SendDate = dmy(SentDate)) %>%
  mutate(Weekday = wday(SentDate,
                        label = TRUE,
                        abbr = FALSE))
```

### Wrangling attributes

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

## Creating network objects using tidygraph

Two functions of **tidygraph** package can be used to create network objects, they are:

-   [`tbl_graph()`](https://tidygraph.data-imaginist.com/reference/tbl_graph.html) creates a **tbl_graph** network object from nodes and edges data.

-   [`as_tbl_graph()`](https://tidygraph.data-imaginist.com/reference/tbl_graph.html) converts network data and objects to a **tbl_graph** network. Below are network data and objects supported by `as_tbl_graph()`

    -   a node data.frame and an edge data.frame,

    -   data.frame, list, matrix from base,

    -   igraph from igraph,

    -   network from network,

    -   dendrogram and hclust from stats,

    -   Node from data.tree,

    -   phylo and evonet from ape, and

    -   graphNEL, graphAM, graphBAM from graph (in Bioconductor).

### The dplyr verbs in tidygraph

-   *activate()* verb from **tidygraph** serves as a switch between tibbles for nodes and edges. All dplyr verbs applied to **tbl_graph** object are applied to the active tibble.
-   In the above the *.N()* function is used to gain access to the node data while manipulating the edge data. Similarly *.E()* will give you the edge data and *.G()* will give you the **tbl_graph** object itself.

### Using tbl_graph() to build tidygraph data model

In this section, you will use `tbl_graph()` of **tinygraph** package to build an tidygraph’s network graph data.frame.

Before typing the codes, you are recommended to review to reference guide of [`tbl_graph()`](https://tidygraph.data-imaginist.com/reference/tbl_graph.html)

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
```

### Reviewing the output tidygraph’s graph object

```{r}
GAStech_graph
```

### Reviewing the output tidygraph’s graph object

-   The output above reveals that *GAStech_graph* is a tbl_graph object with 54 nodes and 4541 edges.

-   The command also prints the first six rows of “Node Data” and the first three of “Edge Data”.

-   It states that the Node Data is **active**. The notion of an active tibble within a tbl_graph object makes it possible to manipulate the data in one tibble at a time.

### Changing the active object

The nodes tibble data frame is activated by default, but you can change which tibble data frame is active with the activate() function. Thus, if we wanted to rearrange the rows in the edges tibble to list those with the highest “weight” first, we could use activate() and then arrange().

```{r}
GAStech_graph %>%
  activate(edges) %>%
  arrange(desc(Weight))
```

Visit the reference guide of [*activate()*](https://tidygraph.data-imaginist.com/reference/activate.html) to find out more about the function.

## Plotting Static Network Graphs with ggraph package

[**ggraph**](https://ggraph.data-imaginist.com/) is an extension of **ggplot2**, making it easier to carry over basic ggplot skills to the design of network graphs.

As in all network graph, there are three main aspects to a **ggraph**’s network graph, they are:

-   [nodes](https://cran.r-project.org/web/packages/ggraph/vignettes/Nodes.html),

-   [edges](https://cran.r-project.org/web/packages/ggraph/vignettes/Edges.html) and

-   [layouts](https://cran.r-project.org/web/packages/ggraph/vignettes/Layouts.html).

### Plotting a basic network graph

The code chunk below uses [*ggraph()*](https://ggraph.data-imaginist.com/reference/ggraph.html), [*geom-edge_link()*](https://ggraph.data-imaginist.com/reference/geom_edge_link.html) and [*geom_node_point()*](https://ggraph.data-imaginist.com/reference/geom_node_point.html) to plot a network graph by using *GAStech_graph*. Before your get started, it is advisable to read their respective reference guide at least once.

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```

### Changing the default network graph theme

In this section, you will use [*theme_graph()*](https://ggraph.data-imaginist.com/reference/theme_graph.html) to remove the x and y axes. Before your get started, it is advisable to read it’s reference guide at least once.

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

### Changing the coloring of the plot

Furthermore, `theme_graph()` makes it easy to change the coloring of the plot.

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes(colour = 'grey50')) +
  geom_node_point(aes(colour = 'grey40'))

g + theme_graph(background = 'grey10',
                text_colour = 'white')
```

### Working with ggraph’s layouts

**ggraph** support many layout for standard used, they are: star, circle, nicely (default), dh, gem, graphopt, grid, mds, spahere, randomly, fr, kk, drl and lgl. Figures below and on the right show layouts supported by `ggraph()`.

### Fruchterman and Reingold layout

The code chunks below will be used to plot the network graph using Fruchterman and Reingold layout.

```{r}
g <- ggraph(GAStech_graph, 
            layout = "fr") +
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

Thing to learn from the code chunk above:

-   *layout* argument is used to define the layout to be used.

### Modifying network nodes

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes()) +
  geom_node_point(aes(colour = Department, 
                      size = 3))

g + theme_graph()
```

-   *geom_node_point* is equivalent in functionality to *geo_point* of **ggplot2**. It allows for simple plotting of nodes in different shapes, colours and sizes. In the codes chnuks above colour and size are used.

### Modifying edges

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") +
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 3)

g + theme_graph()
```

Things to learn from the code chunks above:

-   *geom_edge_link* draws edges in the simplest way - as straight lines between the start and end nodes. But, it can do more that that. In the example above, argument *width* is used to map the width of the line in proportional to the Weight attribute and argument alpha is used to introduce opacity on the line.

## Creating facet graphs

Another very useful feature of **ggraph** is faceting. In visualising network data, this technique can be used to reduce edge over-plotting in a very meaning way by spreading nodes and edges out based on their attributes. In this section, you will learn how to use faceting technique to visualise network data.

There are three functions in ggraph to implement faceting, they are:

-   [*facet_nodes()*](https://r4va.netlify.app/chap27) whereby edges are only draw in a panel if both terminal nodes are present here,

-   [*facet_edges()*](https://ggraph.data-imaginist.com/reference/facet_edges.html) whereby nodes are always drawn in al panels even if the node data contains an attribute named the same as the one used for the edge facetting, and

-   [*facet_graph()*](https://ggraph.data-imaginist.com/reference/facet_graph.html) faceting on two variables simultaneously.

### Working with facet_edges()

In the code chunk below, [*facet_edges()*](https://ggraph.data-imaginist.com/reference/facet_edges.html) is used. Before getting started, it is advisable for you to read it’s reference guide at least once.

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)

g + facet_edges(~Weekday)
```

### Working with facet_edges()

The code chunk below uses theme() to change the position of the legend.

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2) +
  theme(legend.position = 'bottom')
  
g + facet_edges(~Weekday)
```

### A framed facet graph

```{r}
set_graph_style() 

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_edges(~Weekday) +
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```

### Working with facet_nodes()

In the code chunkc below, [*facet_nodes()*](https://ggraph.data-imaginist.com/reference/facet_nodes.html) is used. Before getting started, it is advisable for you to read it’s reference guide at least once.

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_nodes(~Department)+
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```

## Network Metrics Analysis

### Computing centrality indices

```{r}
g <- GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness()) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
            size=betweenness_centrality))
g + theme_graph()
```

Things to learn from the code chunk above:

-   *mutate()* of **dplyr** is used to perform the computation.

-   the algorithm used, on the other hand, is the *centrality_betweenness()* of **tidygraph**.

### Visualising network metrics

It is important to note that from **ggraph v2.0** onward tidygraph algorithms such as centrality measures can be accessed directly in ggraph calls. This means that it is no longer necessary to precompute and store derived node and edge centrality measures on the graph in order to use them in a plot.

```{r}
g <- GAStech_graph %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department, 
                      size = centrality_betweenness()))
g + theme_graph()
```

### Visualising Community

```{r}
g <- GAStech_graph %>%
  mutate(community = as.factor(
    group_edge_betweenness(
      weights = Weight, 
      directed = TRUE))) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(
    aes(
      width=Weight), 
    alpha=0.2) +
  scale_edge_width(
    range = c(0.1, 5)) +
  geom_node_point(
    aes(colour = community))  

g + theme_graph()
```

In order to support effective visual investigation, the community network above has been revised by using [`geom_mark_hull()`](https://ggforce.data-imaginist.com/reference/geom_mark_hull.html) of [ggforce](https://ggforce.data-imaginist.com/) package.

```{r}
g <- GAStech_graph %>%
  activate(nodes) %>%
  mutate(community = as.factor(
    group_optimal(weights = Weight)),
         betweenness_measure = centrality_betweenness()) %>%
  ggraph(layout = "fr") +
  geom_mark_hull(
    aes(x, y, 
        group = community, 
        fill = community),  
    alpha = 0.2,  
    expand = unit(0.3, "cm"),  # Expand
    radius = unit(0.3, "cm")  # Smoothness
  ) + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(fill = Department,
                      size = betweenness_measure),
                      color = "black",
                      shape = 21)
  
g + theme_graph()
```

## Building Interactive Network Graph with visNetwork

-   [visNetwork()](http://datastorm-open.github.io/visNetwork/) is a R package for network visualization, using [vis.js](http://visjs.org/) javascript library.

-   *visNetwork()* function uses a nodes list and edges list to create an interactive graph.

    -   The nodes list must include an “id” column, and the edge list must have “from” and “to” columns.

    -   The function also plots the labels for the nodes, using the names of the actors from the “label” column in the node list.

-   The resulting graph is fun to play around with.

    -   You can move the nodes and the graph will use an algorithm to keep the nodes properly spaced.

    -   You can also zoom in and out on the plot and move it around to re-center it.



### Data preparation

Before we can plot the interactive network graph, we need to prepare the data model by using the code chunk below.

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  left_join(GAStech_nodes, by = c("sourceLabel" = "label")) %>%
  rename(from = id) %>%
  left_join(GAStech_nodes, by = c("targetLabel" = "label")) %>%
  rename(to = id) %>%
  filter(MainSubject == "Work related") %>%
  group_by(from, to) %>%
    summarise(weight = n()) %>%
  filter(from!=to) %>%
  filter(weight > 1) %>%
  ungroup()
```


### Plotting the first interactive network graph

The code chunk below will be used to plot an interactive network graph by using the data prepared.

```{r}
visNetwork(GAStech_nodes, 
           GAStech_edges_aggregated)
```

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") 
```


### Working with visual attributes - Nodes

```{r}
GAStech_nodes <- GAStech_nodes %>%
  rename(group = Department) 
```

When we rerun the code chunk below, visNetwork shades the nodes by assigning unique colour to each category in the group field.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

### Working with visual attributes - Edges

In the code run below visEdges() is used to symbolise the edges.
- The argument arrows is used to define where to place the arrow.
- The smooth argument is used to plot the edges using a smooth curve.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visEdges(arrows = "to", 
           smooth = list(enabled = TRUE, 
                         type = "curvedCW")) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```


### Interactivity

In the code chunk below, visOptions() is used to incorporate interactivity features in the data visualisation.

The argument highlightNearest highlights nearest when clicking a node.
The argument nodesIdSelection adds an id node selection creating an HTML select element.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = TRUE) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

